---
title: "Intro to R for Analytics"
author: "Kevin Terwilliger"
date: "11/20/2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(dplyr)
library(MASS)

set.seed(1)
```

## An Introduction to R

R is a language and environment for statistical computing and graphics.

It provides a wide variety of statistical (linear and nonlinear modelling, classical statistical tests, time-series analysis, classification, clustering, …) and graphical techniques, as well as helpful tools for publication.

## Installing R

Visit this link: https://cloud.r-project.org/ and download the file for your correct operating system.

Rstudio is a fantastic, free software for programming in R. You can download it here:

https://rstudio.com/products/rstudio/download/?utm_source=downloadrstudio&utm_medium=Site&utm_campaign=home-hero-cta#download

## R basics

```{r, echo = TRUE}
data(mtcars)
names(mtcars)
head(mtcars)
```

## Math and Symbols

```{r echo=T}
a <- 8
## Sample 1 value between 1 and 30
b = sample(c(1:30),1)
paste(a,b,sep = ", ")
a + b
a*b
a^b
a/b
a == b
```


## Vectors and Lists

```{r echo=T}
v1 = c(1,2,3,4)
v2 = c(1:4)
## Check if vectors hold the same values
v1 == v2
```

```{r echo=T}
## We can make a list of objects with
list = list("Column1" = c(1:5),
            "Column2" = sample(c(1:20),size = 5,replace = F))
list
as.data.frame(list)

## This won't work
# list = list("Column1" = c(1:5),  
#             "Column2" = sample(c(1:20),size = 6,replace = F))
# as.data.frame(list)
```


## Graphics

```{r echo=T}
## Randomly generate 100 values from a normal distribution (mean=0,sd=1)
x = rnorm(100)
y = rnorm(100)
plot(x,y,xlab="this is the x-axis",ylab="this is the y-axis",main="Plot of X vs Y")
```

### Data Frames and Indexing

```{r}
names(Boston)
```

