 ---
title: "An Introduction to R"
author: "Kevin Terwilliger"
date: "11/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(dplyr)
library(MASS)

set.seed(1)
```

R is a language and environment for statistical computing and graphics.

It provides a wide variety of statistical (linear and nonlinear modelling, classical statistical tests, time-series analysis, classification, clustering, …) and graphical techniques, as well as helpful tools for publication.

## Installing R

Visit this link: https://cloud.r-project.org/ and download the correct file for your operating system (Windows,MacOS,Linux).

Rstudio is a fantastic, free software for programming in R. You can download it here:

https://rstudio.com/products/rstudio/download/?utm_source=downloadrstudio&utm_medium=Site&utm_campaign=home-hero-cta#download

## R basics

```{r, echo = TRUE}
data(mtcars)
names(mtcars)
head(mtcars)
```

## Math and Symbols

```{r echo=T}
a <- sample(c(1:30),1)
## Sample 1 value between 1 and 30
b = sample(c(1:30),1)
paste(a,b,sep = ", ")
a + b
a*b
a^b
a/b
a == b
```


## Vectors and Lists

```{r echo=T}
v1 = c(1,2,3,4)
v2 = c(1:4)
## Check if vectors hold the same values
v1 == v2
```

```{r echo=T}
## We can make a list of objects with
list = list("Column1" = c(1:5),
            "Column2" = sample(c(1:20),size = 5,replace = F))
list
as.data.frame(list)

## This won't work
# list = list("Column1" = c(1:5),
#             "Column2" = sample(c(1:20),size = 6,replace = F))
# as.data.frame(list)
```

## Graphics

```{r echo=T}
## Randomly generate 100 values from a normal distribution (mean=0,sd=1)
x = rnorm(100)
y = rnorm(100)
plot(x,y,xlab="this is the x-axis",ylab="this is the y-axis",main="Plot of X vs Y")
```

### Data Frames and Indexing

```{r echo=T}
Insurance = Insurance
nrow(Insurance)
ncol(Insurance)
## OR
dim(Insurance)
## Print Column Names
names(Insurance)
## Same as head(Insurance)
Insurance[1:5,]

Insurance[1:5,c("Age","Holders","Claims")]

class(Insurance$Holders)

Insurance[1:5,-c(1:(length(Insurance)-2))]
```

## Writing/Reading Data

```{r echo=T}
write.csv(cars,file="~/Desktop/cars.csv")
setwd("~/Desktop/GitHub/Econ490/Report4/")
getwd()
songs = read.csv("song_data.csv",
                 header = T)
```

## Descriptive Statistics and Regressions

```{r echo=T}
summary(Boston)

train = sample_frac(Boston,0.7)
test = setdiff(Boston,train)

## Single Linear Regression
fit = lm(medv ~ lstat,data = train)
summary(fit)

plot(Boston$lstat,Boston$medv) +
  abline(fit,lwd=3,col="red")
```

```{r echo=T}
predictions = predict(fit,test)

ggplot() +
  geom_point(aes(x = test$medv, y = predictions),
             colour = 'blue') +
  geom_line(aes(x = test$medv, y = test$medv),
            colour = 'red') + # plotting Y againts itself produces a red 45 degree line 
  xlab('actual y') + ylab('predicted y')
```

```{r}
## Multiple Linear Regression
fit.2var = lm(medv ~ age+lstat,data=train)
summary(fit.2var)

fit.allVars = lm(medv~.,data=train)
summary(fit.allVars)

pred.2var = predict(fit.2var,test)
pred.all = predict(fit.allVars,test)

df = data.frame(x = c(1:length(pred.2var)),
                "2var" = pred.2var,
                "all" = pred.all)

head(df)

library(reshape2)
library(dplyr)

melted.df = melt(df,id.vars="x",
                 variable.name = "Model",
                 value.name = "Predictions")

melted.df$Model = recode(melted.df$Model,
                         X2var = "2 Variable Model",
                         all = "All Variables Model")

head(melted.df)

ggplot(melted.df,aes(x=rep(test$medv,2),y=Predictions,
                     shape = Model, 
                     color = Model)) +
  geom_point() +
  geom_point() +
  geom_line(aes(y = rep(test$medv,2)),
            colour = 'red') +
  xlab('actual y') + ylab('predicted y')
```

